Covariate analysis of CMC RNASeq data reprocessed with STAR-feature counts workflow with Ensembl v75 (ACC and DLPFC with CQN + SVA normlaisation)

### 2018-02-19 code/Rmd/CMC_ACC_DLPFC_STAR_ENSEMBL_v75_CQN_SVA.Rmd
Minor modifications to ACC and DLPFC processing markdown with SVA

---
title: "Covariate analysis of CMC RNASeq data reprocessed with STAR-feature counts workflow with Ensembl v75 (ACC and DLPFC with CQN normlaisation)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse)

@@

library(ggplot2)
library(reshape2)
library(limma)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)
library(RColorBrewer)
library(cqn)
library(sva)

@@

@@

# Source modified functions from sva package
source('../R/irwsva.build.R')
source('../R/f.pvalue.R')

# Compute actual variance of all principal components
tmp = svd(RESIDUAL.GENE_EXPRESSION)
actual.var = tmp$d^2/sum(tmp$d^2)

# Compute permuted variance of all principal components
permuted.var = foreach(i = 1:20, .combine = rbind, .packages = c('dplyr')) %dopar% {
  tmp.residual = t(apply(RESIDUAL.GENE_EXPRESSION, 1, sample, replace = FALSE))
  tmp = svd(tmp.residual)
  permuted.var = tmp$d^2/sum(tmp$d^2)
}
permuted.var = apply(permuted.var, 2, mean)
NUM.SV = sum(actual.var >= permuted.var[1])

@@

# Estimate surrogate variables
SURR.VAR = irwsva.build(CQN.GENE_EXPRESSION$E, MODEL1, MODEL0, n.sv = NUM.SV, B = 30, tol = 1e-20)$sv
colnames(SURR.VAR) = paste0('SV',1:dim(SURR.VAR)[2])
rownames(SURR.VAR) = rownames(MODEL1)

# Estimate voom weights for dispersion control
VOOM.SVA.WEIGHTS = voom(NEW.COUNTS, 
                        design=cbind(MODEL1, SURR.VAR), plot=F)
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E

# Calculate correlation values of random effects
correlation = parallelDuplicateCorrelation(VOOM.SVA.WEIGHTS, block = COVARIATES$Individual_ID)

# Re-estimate voom weights
VOOM.SVA.WEIGHTS = voom(NEW.COUNTS, 
                        design = cbind(MODEL1, SURR.VAR), plot=F,
                        block = COVARIATES$Individual_ID, 
                        correlation = correlation$cor)
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E

@@

# Add variable of interest back to the residuals
varsToAddIn = grep("Dx.Tissue", colnames(DESIGN), value = T)
RESIDUAL.SVA.GENE_EXPRESSION = RESIDUAL.SVA.GENE_EXPRESSION + 
  VOOM.SVA.FIT$coefficients[,varsToAddIn] %*% t(VOOM.SVA.FIT$design[,varsToAddIn])
```
Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
cr = cor(t(RESIDUAL.SVA.GENE_EXPRESSION))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Clustering residual data
```{r decompse.normalise.data2, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.SVA.GENE_EXPRESSION, scale.=T, center = T)

@@

tree = hclust(as.dist(t(RESIDUAL.SVA.GENE_EXPRESSION)))
cols = WGCNA::labels2colors(COVARIATES.tmp);

>>>>

---
title: Covariate analysis of CMC RNASeq data reprocessed with STAR-feature counts workflow with Ensembl v75 (ACC and DLPFC with CQN + SVA normlaisation)
author: Thanneer Perumal
editor_options:
  chunk_output_type: console
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse)

@@

library(ggplot2)
library(reshape2)
library(limma)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)
library(RColorBrewer)
library(cqn)

@@

# Source modified utility 
source('../R/parallelDuplicateCorrelation.R')
source('../R/irwsva.build.R')
source('../R/f.pvalue.R')

@@

# Compute actual variance of all principal components
expr = RESIDUAL.GENE_EXPRESSION; expr[is.na(expr)] = 0
tmp = svd(expr)
actual.var = tmp$d^2/sum(tmp$d^2)

# Compute permuted variance of all principal components
permuted.var = foreach(i = 1:20, .combine = rbind, .packages = c('dplyr')) %dopar% {
  tmp.residual = t(apply(expr, 1, sample, replace = FALSE))
  tmp = svd(tmp.residual)
  permuted.var = tmp$d^2/sum(tmp$d^2)
}
permuted.var = apply(permuted.var, 2, mean)
NUM.SV = sum(actual.var >= permuted.var[1])

@@

# Estimate surrogate variables
SURR.VAR = irwsva.build(CQN.GENE_EXPRESSION$E.no.na, MODEL1, MODEL0, n.sv = NUM.SV, B = 30, tol = 1e-20)$sv
colnames(SURR.VAR) = paste0('SV',1:dim(SURR.VAR)[2])
rownames(SURR.VAR) = rownames(MODEL1)

# Estimate voom weights for dispersion control
VOOM.SVA.WEIGHTS = voom(cnts, 
                        design=cbind(MODEL1, SURR.VAR), plot=F)
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E

# Calculate correlation values of random effects
correlation = parallelDuplicateCorrelation(VOOM.SVA.WEIGHTS, 
                                           block = COVARIATES$Individual_ID,
                                           method = 'lmer')

# Re-estimate voom weights
VOOM.SVA.WEIGHTS = voom(cnts, 
                        design = cbind(MODEL1, SURR.VAR), plot=F,
                        block = COVARIATES$Individual_ID, 
                        correlation = correlation$cor)
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E

@@

# Add variable of interest back to the residuals
varsToAddIn = grep("Dx.Tissue", colnames(DESIGN), value = T)
RESIDUAL.SVA.GENE_EXPRESSION = RESIDUAL.SVA.GENE_EXPRESSION + 
  VOOM.SVA.FIT$coefficients[,varsToAddIn] %*% t(VOOM.SVA.FIT$design[,varsToAddIn])
```
Coexpression of genes 
```{r coexp2, cache=FALSE, fig.height=5, fig.width=5}
expr = RESIDUAL.SVA.GENE_EXPRESSION;
expr[is.na(expr)] = 0
cr = cor(t(expr))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

### Clustering residual data
```{r decompse.normalise.data2, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(expr, scale.=T, center = T)

@@

tree = hclust(as.dist(t(expr)))
cols = WGCNA::labels2colors(COVARIATES.tmp);
